Project Report\\\textsc{DEVS Ex Machina}

### 2015-05-18 manual/projectrapport.tex
* added free function for extracting the payload from the message * expanded the part on the messaging size & logger in the report
Signed-off-by: Lttlemoi <lttlemoi+git@gmail.com>

\section{Messaging Size}
A message is constructed once, and after that never copied. Messages are sent and received by reference to smart pointer, incurring a size cost of (on 64-bit platforms) 128bit. The payload (and its size) is irrelevant to the simulation up to the point where the message is actually received. A default implementation allows the payload stored as a std::string, but this can be overridden by way of templated code. %TODO Stijn.

@@

\section{Scheduler}
The scheduler from the prototypes was reused, with little to no changes (barring a few bugfixes). An unsynchronized version was implemented for comparison purposes.

@@

@@

It therefore lacks certain key features such as
	\item crash safe
	\item custom assert that will log an error before shutting down

@@

Chances are that we will wait with implementing these features until we actually need them.
\chapter{Performance Analysis}
Whenever we made big design/architecture decisions, we first wrote some prototyping testcode to prove the viability (e.g. Network, Tracing, Logger).
\section{First comparison}
We use perf to get statistically sound benchmarks. The command run was :
\lstinline!$perf stat -r {rounds} {command}!\\

>>>>

\section{Messaging Size}
A message is constructed once, and after that never copied. Messages are sent and received by reference to smart pointer, incurring a size cost of (on 64-bit platforms) 128bit. The payload (and its size) is irrelevant to the simulation up to the point where the message is actually received. A default implementation allows the payload stored as a std::string, but a specialized message class can be used for sending objects of a different type. When constructing a message by using the functionality provided by the Port class, this happens automatically. When receiving messages, the free function template is provided to extract the object from the message.\\
For example:
\begin{lstlisting}[language=C++,basicstyle=\small\color{black}]
//initialization of necessary objects is left out.
double value;
n_model::t_portptr port;

//create messages that contain a double
auto msgVector = port->createMessages(value);

//say we receive this message somewhere else
n_network::t_msgptr received =  msgVector[0];

//retrieve the data when receiving the message
const double& data = n_network::getMsgPayload<double>(received);

//Assuming that no Z function was applied to the message.
assert(value == data);
\end{lstlisting}

@@

\section{Scheduler}
A Scheduler was implemented based on the prototyping code. As required, as list based scheduler was implemented. \\
The interface from PyPDevs' scheduler class was reused, but internally the scheduler class in our project differs on a few point. \\
\subsection{Composition}
All schedulers are based on a heap, which order items to be scheduled. We never actually schedule models, only entries (A struct containing name and time to schedule), allowing for far faster operation. This is complemented by a hashmap, so that entries can have O(1) lookup and erasure (required for the use case of antimessaging, among others). The operation massReschedule() was left out of the interface, since it can be easily composed using erase() and push\textunderscore back().
\\
\subsection{Implementation}
The actual heap used is either based on std::list as required, or one of Boost's heaps. The user is provided with a convenience factory function that creates Scheduler instances with varying heap backends. In general, the Fibonacci heap is one of the fastest, with skew heap sometimes coming out ahead. The list based heap is always a factor 2 slower, though again this depends on the usage scenario. \\Important to note is that only node based heaps can be used efficiently (without compromising the performance of erase() and contains()). We store on insertion an iterator with the key of the pushed element in the hashmap. This also works as a safety check, preventing double insertion (depending ofcourse on the hash-specialization semantics). This freedom of implementation is visible in the tracers where duplicate elements are allowed if they have a different address, contrasted with the model scheduler which disallows a model to be scheduled twice.\\
Finally, the std::hash$<>$ specialization used by the message scheduler allows O(1) annihilation of a message by its antimessage (since the hash is identical).
\subsection{Differences}
We intended to use the scheduler class not only for scheduling models, but also to store pending incoming messages, and tracing messages. To allow this, the Scheduler class only requires from a scheduled item that it has operator$<$ and std::hash$<$T$>$ defined.

@@

	\item crash safety. The logger will catch the following signals and flush all remaining output before exiting the program:
			\item \emph{SIGSEGV} Segmentation fault, sent by the operating system.
			\item \emph{SIGTERM} Termination request, sent by the program.
			\item \emph{SIGABRT} Abort --- abnormal termination, e.g. when an assertion fails.
			\item \emph{SIGINT} Interrupt, used for catching crtl+C, which is useful for debugging deadlocks.

@@

@@

\chapter{Performance Analysis}
Whenever we made big design/architecture decisions, we first wrote some prototyping testcode to prove the viability (e.g. Network, Tracing, Logger).
\section{First comparison}
We use perf to get statistically sound benchmarks. The command run was :
\lstinline!$perf stat -r {rounds} {command}!\\
